---
title: "p8105_hw2_yy3295"
author: "Ye Yuan"
date: "2023-10-04"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```

##Problem 1
We clean the 538 `pols` data, which provides information on the number of national politicians who are democratic or republican at any given time. There are some values for which `prez_gop` is `2` -- these are months in which Ford became President following Nixon's resignation. In the new `president` variable created as part of our data cleaning, we code these as `gop` (same as values when `prez_gop` is `1`).

```{r clean_538_pols}
month_df = 
  tibble(
    month_num = 1:12,
    month_abb = month.abb,
    month = month.name
  )

pols = 
  read_csv("data/pols-month.csv") |>
  separate(mon, into = c("year", "month_num", "day"), convert = TRUE) |>
  mutate(
    president = recode(prez_gop, "0" = "dem", "1" = "gop", "2" = "gop")) |>
  left_join(x = _, y = month_df) |> 
  select(year, month, everything(), -day, -starts_with("prez")) 
```

We also clean the 538 `snp` data, which contains information related to Standard & Poor’s stock market index.

```{r clean_538_snp}
snp = 
  read_csv(
    "data/snp.csv",
    col_types = cols(date = col_date(format = "%m/%d/%y"))) |>
  separate(date, into = c("year", "month_num", "day"), convert = TRUE) |>
  mutate(
    year = if_else(year > 2023, year - 100, year)) |> 
  left_join(x = _, y = month_df) |> 
  select(year, month, close) 
```

Finally, we tidy the `unemployment` data so that it can be merged with the `pols` and `snp` datasets.

```{r clean_538_unemp}
unemployment = 
  read_csv("data/unemployment.csv") |>
  rename(year = Year) |>
  pivot_longer(
    Jan:Dec, 
    names_to = "month_abb",
    values_to = "unemployment"
  ) |> 
  left_join(x = _, y = month_df) |> 
  select(year, month, unemployment)
```

Now we merge the three datasets!

```{r merge_538}
data_538 = 
  left_join(pols, snp) |>
  left_join(x = _, y = unemployment)

str(data_538)
```

Notice that there are some `NA` values in the `close` and `unemployment` variables, which indicate that the value of these variables is missing at those locations.

Let's talk about the 538 datasets. The `pols` data has `r nrow(pols)` observations and `r ncol(pols)` variables and tells us about the party affiliation distribution (democrat or republican) for governors and senators for a given year from years `r pols |> pull(year) |> min()` to `r pols |> pull(year) |> max()`. It also tells us whether the sitting president was a democrat or republican. The `snp` data has `r nrow(snp)` observations and `r ncol(snp)` variables, ranging from years `r snp |> pull(year) |> min()` to `r snp |> pull(year) |> max()`. The `unemployment` data has `r nrow(unemployment)` observations and `r ncol(unemployment)` variables ranging from years `r unemployment |> pull(year) |> min()` to `r unemployment |> pull(year) |> max()`. In Januarys in or after 1975 in which a democrat was president, the **average unemployment rate was `r filter(data_538, month == "January", year >= 1975, president == "dem") |> pull(unemployment) |> mean() |> round(2)`**.  The average unemployment rate over the same time period in which a republican was president was `r filter(data_538, month == "January", year >= 1975, president == "gop") |> pull(unemployment) |> mean() |> round(2)`.

#Problem 2
Read and clean the Mr. Trash Wheel sheet:
specify the sheet in the Excel file and to omit non-data entries using arguments in read_excel
use reasonable variable names
omit rows that do not include dumpster-specific data

```{r}
Mr_TW_df = 
  read_excel("data/202309 Trash Wheel Collection Data.xlsx",
             sheet = "Mr. Trash Wheel", range = "A2:N586") |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(homes_powered = weight_tons * 500 / 30,
         wheel_type = c("MTW"),
         year = as.numeric(year)) |>
  relocate(dumpster, wheel_type)

Mr_TW_df
```

Import, clean, and organize the data for Professor Trash Wheel

```{r}
Prof_TW_df = 
  read_excel("data/202309 Trash Wheel Collection Data.xlsx",
  sheet = "Professor Trash Wheel", range = "A2:M108") |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(homes_powered = weight_tons * 500 / 30,
         wheel_type = c("PTW")) |>
  relocate(dumpster, wheel_type)

Prof_TW_df
```

Import, clean, and organize the data for Gwynnda Trash Wheel

```{r}
Gwynnda_TW_df = 
  read_excel("data/202309 Trash Wheel Collection Data.xlsx",
  sheet = "Gwynnda Trash Wheel", range = "A2:L157") |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(homes_powered = weight_tons * 500 / 30,
         wheel_type = c("GTW")) |>
  relocate(dumpster, wheel_type)

Gwynnda_TW_df
```

Combine the Professor and Gwynnda Trash Wheel with the Mr. Trash Wheel dataset to produce a single tidy dataset

```{r}
Final_TW_df = bind_rows(Mr_TW_df, Prof_TW_df, Gwynnda_TW_df)

Final_TW_df
```
Write a paragraph about these data.

In Mr.Trash Wheel data set, there are 584 rows and 16 cols. The key variables are wheel_type, dumpster, weight_tons, homes_powered.
In Professor Trash Wheel data set, there are 106 rows and 14 cols. The key variables are wheel_type, dumpster, weight_tons, homes_powered.
In Gwynnda Trash Wheel data set, there are 155 rows and 13 cols. The key variables are wheel_type, dumpster, weight_tons, homes_powered.
In the final combined Trash Wheel data set, there are 845 rows and 15 cols. The key variables are wheel_type, dumpster, weight_tons, homes_powered.
The total weight of trash collected by Professor Trash Wheel were 216.26 tons.

```{r}
filter(Final_TW_df, wheel_type == "PTW") |>
  pull(weight_tons) |>
  sum()
```

The total number of cigarette butts collected by Gwynnda in July of 2021 are 16300.
```{r}
filter(Final_TW_df, wheel_type == "GTW", year == "2021", month == "July") |>
  pull(cigarette_butts) |>
  sum()
```

#Problem 3
```{r}
MCI_Base = 
  read_csv("data/MCI_baseline.csv") |>
  janitor::clean_names() |>
  drop_na(age_at_onset) |>
  mutate()

MCI_Base
```

```{r}
MCI_Amy = 
  read_csv("data/MCI_amyloid.csv") |>
  janitor::clean_names() |>
  pivot_longer() |>
  mutate(
    years_since_diagnosed = replace(years_since_diagnosed, years_since_diagnosed == "bl", "0"),
    years_since_diagnosed = as.numeric(years_since_diagnosed))

MCI_Amy
```

```{r}

```


